### 2021-06-24 formerPapers/grlRrevision2/pointbypoint/point-by-point.tex

\documentclass[11pt]{article}
% General document formatting
\usepackage[margin=0.75in]{geometry}
\usepackage[parfill]{parskip}
\usepackage[utf8]{inputenc}
\usepackage{subfig}         % side-by-side figures 
% Related to math
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{titling}
\usepackage{hyperref}
\usepackage{wrapfig}
\usepackage{booktabs} % for wrapping tabulars in accord with
\bibliographystyle{agu}

%\usepackage[math]{kurier}
\newcommand\be{\begin{equation}} % shortcut to start eq envs 
\newcommand\ee{\end{equation}}   % shortcut to end eq envs
\newcommand\ol{\overline}        % shortcut to draw overline 
\newcommand\bra{\langle}
\newcommand\ket{\rangle}
\newcommand\El{\mathcal{L}}
\newcommand\tg{\tilde{g}}
\newcommand\tG{\tilde{G}}
\begin{document}

	\title{Response to reviewers: ``Back to Einstein: burial-induced three-range diffusion in fluvial sediment transport''}
	\author{J. Kevin Pierce \& Marwan A. Hassan}
	\maketitle

\section*{Associate Editor}
I agree with the reviewers that the revisions have improved the manuscript but there are new points raised by the reviewers, in particular the comments of reviewer 1 on the number of regimes identified and the need for a consistent story, rather than a timeline style of presentation of the results, and edits from reviewer 2.

\paragraph{Reply:}
We modified the manuscript in response to every comment of both reviewers. 
We clarified the number of regimes identified (four variance scaling ranges, three of which are diffusive) at L20, 202, 226, and 273. We removed the timeline style presentation of the results, and we incorated all edits from reviewer \#2.

\section*{Reviewer \#1}

\begin{enumerate}
\item 'However, this context changes the main story and we now have a manuscript with “three range diffusion” in the title, but “four range” in the key points and the key figure- figure 2. Also, lines 221-226 describe three ranges, then the 4 th range is mentioned after the list but not counted.'
\paragraph{Reply:} 
We reconciled the inconsistency you pointed out between "three range" and "four range" within the revised manuscript. There are four variance scaling ranges but only three ``diffusion'' ranges. The fourth ``geomorphic'' variance scaling range does not involve the movement of any particles (at least with permanent burial), so it isn't a diffusion range. We clarified this distinction in the revised manuscript at L20, 202, 226, and 273 so we could write ``three-range diffusion'' without ambiguity.

\item I suggest that the authors remove their timeline of discovery (eg lines 240-250 “in the preliminary studies for this paper we found that a two-state....” “these realizations and the need for a physical model...” )
\paragraph{Reply:}
We removed the timeline of discovery and rephrased the underlying point that the Nikora et al. suggestion of a two state model with heavy-tailed rests does not generate three diffusion ranges.

\item  All previous descriptions of the global regime fit within this model depending on the timescale/tail relationship between rest and burial. Similarly, existence of a geomorphic range (or at least the slope on figure 2) is a function of permanent burial or not. The multi state model generalizes all random walk findings and highlights the physical process timescales that are significant. If there are other physical processes that are relevant, they can be
added. All of this should be reflected in Figure 2-
\paragraph{Reply:} 
We comprehensively revised figure 2 to include all components of your thoughtful summary. We highlighted the variation of global range diffusion with $k_1/\kappa$ (ratio of burial time to resting time), we illustrated the anticipated result of impermanent burial in the geomorphic range, and we added an additional panel with a schematic of bedload trajectories through motion, rest, and burial states to indicate the physical processes that generate each diffusion range.
\end{enumerate}

We appreciate your attention toward improving our manuscript and hope we have addressed your concerns sufficiently.

\section*{Reviewer \#2}

\subsection*{General Comments}

The revised manuscript appropriately addresses the first set of reviews by placing a larger emphasis on the geophysical aspects of the problem. The revised manuscript is suitable for publication following minor revisions or a few changes that could potentially happen within the copy-editing phase.

\subsection*{Specific Comments:}
\begin{enumerate}
\item Abstract. The abstract would benefit from some general editing. 7/12 lines are introducing the problem rather than serving as a synopsis of the paper. Some of this would be better suited to the introduction section.
\paragraph{Reply:} We removed these lines and replaced them with a synopsis of the paper. The first several lines of the abstract now introduce the problem, the middle lines state what we did to solve the problem, and the final lines summarize our key findings.

\item Ln. 64. This is a bit of a miss citaiton here. Both Bradley (2017) and Phillips et al. (2013) only consider time above a transport threshold thus specifically excluding the time spent resting during low flow between floods.
\paragraph{Reply:} We modified this sentence from the terminology ``between subsequent floods'' to represent our point that sampling timescales of tracer positions in these studies are at least as long as individual floods.

\item Ln. 278-282. In Weeks and Swinney (1998) long time subdiffusion is uncommon for asymmetric random walks requiring very heavy-tailed waiting times. Given re-entrainment from burial is likely in all but aggrading systems is long term subdiffusion really a viable solution for natural rivers?
\paragraph{Reply:} We are unsure whether long-time sub-diffusion is viable, so we modified this section to better communicate the existing support for both possibilities. The field study of  \citet{Olinde2015} shows sufficiently heavy-tailed resting times for long-time sub-diffusion, but their methods do not discern whether these rests are due to burial or some other process. Several models including our recent JGR paper show that burial times under equilibrium transport conditions are not sufficiently heavy-tailed for sub-diffusion \citep{Pierce2020,Martin2014,Voepel2013}. Similarly, \citet{Bradley2017} presents resting times that were insufficient for long-time sub-diffusion. The evidence mostly points toward long-time super-diffusion, but the limited data leave us not confident enough to refute the Nikora et al. hypothesis.
\end{enumerate}

Thank you sincerely for your constructive comments and attention toward our work.

\bibliography{biblio}
\end{document}

### 2021-06-24 abs.tex
got one chapter in. some issues.

\chapter{Sediment transport and landscape evolution}
\label{ch:Introduction}

Landscapes evolve across geologial time due to energy and mass contributed from above and below the Earth's surface.

\section{Stochastic properties of sediment movement}

\section{Historical survey: statistical physics in sediment transport}

\section{Philosophy of the statistical physics approach}

\section{What's left?}

\endinput

Any text after an \endinput is ignored.
You could put scraps here or things in progress.

>>>>

\chapter{Abstract}

Write an abstract.

### 2021-06-24 formerPapers/grlRrevision2/rwdiffu.tex
got one chapter in. some issues.

%%!TEX root = diss.tex

\chapter{Inclusion of sediment burial}
\label{ch:downDiff}

@@

\section{Bedload trajectories as a multi-state random walk}
\label{sec:model}
\subsection{Model assumptions}
\label{sec:assumptions}
We construct a three-state random walk where the states are motion, surface rest, and burial, and we label these states as $i=2$ (motion), $i=1$ (rest), and $i=0$ (burial).
Our target is the probability distribution $p(x,t)$ to find a grain at position $x$ and time $t$ if we know it started with the initial distribution $p(x,0)=\delta(x)$.
We characterize times spent moving or resting on the surface by exponential distributions $\psi_2(t)=k_2e^{-k_2 t}$ and $\psi_1(t) = k_1e^{-k_1t}$, since numerous experiments show thin-tailed distributions for these quantities \citep{Fathel2015,Roseberry2012,Einstein1937,Ancey2006,Martin2012}. We expect our conclusions will not be contingent on the specific distributions chosen, since all thin-tailed distributions provide similar diffusion characteristics in random walks \citep{Weiss1994,Weeks1998}.
We consider grains in motion to have characteristic velocity $v$ \citep{Lisle1998,Lajeunesse2018}, and we model burial as long lasting enough to be effectively permanent \citep{Wu2019}, with grains resting on the surface having a probability per unit time $\kappa$ to become buried, meaning $\Phi(t) = e^{-\kappa t}$ represents the probability that a grain is not buried after resting for a time $t$, while $1-\Phi(t)$ represents the probability that it is buried.
We specify the initial conditions with probabilities $\theta_1$ and $\theta_2$ to be in rest and motion at $t=0$, and we require $\theta_1+\theta_2=1$ for normalization.

@@

\begin{figure}[t]
	\includegraphics[width=\linewidth,keepaspectratio]{./figures/ch4/pdf-plot.png}
	\caption{Joint distributions for a grain to be at position $x$ at time $t$ are displayed for the choice $k_1=0.1$, $k_2=1.0$, $v=2.0$. Grains are considered initially at rest ($\theta_1=1$, $\theta_2=0$). The solid lines are the analytical distribution in equation (\ref{eq:pdf}), while the points are numerically simulated, showing the correctness of our derivations. Colors pertain to different times. Units are unspecified, since we aim to demonstrate the general characteristics of $p(x,t)$. Panel (a) shows the case $\kappa=0$ -- no burial. In this case, the joint distribution tends toward Gaussian at large times \citep{Einstein1937,Lisle1998}. Panel (b) shows the case when grains have rate $\kappa = 0.01$ to become buried while resting. Because of burial, the joint distribution tends toward a more uniform distribution than Gaussian.
		\label{fig:pdfs}}
\end{figure}

@@

\begin{figure}[h!]	
	\includegraphics[width=\linewidth,keepaspectratio]{./figures/ch4/diffusion.png}
	\caption{Panel (a) sketches conceptual trajectories of three grains, while panel (b) depicts the variance (\ref{eq:var}) with mean motion time $1.5$ s, resting time $30.0$ s, and movement velocity $0.1$ m/s -- values comparable to laboratory experiments transporting small ($5$ mm) gravels \citep{Lajeunesse2010,Martin2012}. The burial timescale is $7200.0$s (two hours), and grains start from rest ($\theta_1=1$). The solid line is equation (\ref{eq:var}), and the points are numerically simulated. Panel (b) demonstrates four distinct scaling ranges of $\sigma_x^2$: local, intermediate, global, and geomorphic. The first three are diffusive. Three crossover times $\tau_L$, $\tau_I$, and $\tau_G$ divide the ranges. Within each range, a slope key demonstrates the scaling $\sigma_x^2 \propto t^\gamma$. Panel (a) demonstrates that different mixtures of motion, rest, and burial states generate the ranges. At local timescales, grains usually either rest or move; at intermediate timescales, they transition between rest and motion; at global timescales, they transition between rest, motion, and burial; and at geomorphic timescales, all grains bury. Additional slope keys in the local and global ranges of panel (b) illustrate the effect of initial conditions and rest/burial timescales on the diffusion, while the additional slope key within the geomorphic range demonstrates the expected scaling when burial is not permanent, as we discuss in section \ref{sec:discussion}.
	}
	\label{fig:var}
\end{figure}

@@

The local and intermediate range diffusion characteristics resulting from our model correspond closely to the original Nikora et al. concepts, while our global range has a different origin than Nikora et al. envisioned.
\citet{Nikora2001a} explained that local diffusion results from the non-fractal (smooth) characteristics of bedload trajectories between subsequent interactions with the bed,  while intermediate diffusion results from the fractal (rough) characteristics of bedload trajectories after many interactions with the bed.
Our model represents these conclusions: non-fractal (and super-diffusive) bedload trajectories exist on timescales short enough that each grain is either resting or moving, while fractal (and normal-diffusive) bedload trajectories exist on timescales when grains are actively switching between motion and rest states.
We conclude that local and intermediate ranges stem from the interplay between motion and rest timescales, as demonstrated by earlier two-state random walk models \citep{Lisle1998,Lajeunesse2017} and by all Newtonian models that develop sequences of motions and rests \citep{Nikora2001a, Bialik2012}, even those including heavy-tailed rests \citep{Fan2016}.

@@

The analytical solution of bedload diffusion in equation (\ref{eq:var}) reduces exactly to the analytical solutions of the \citet{Lisle1998} and \citet{Lajeunesse2017} models in the limit without burial ($\kappa \rightarrow 0$), the \citet{Wu2019} model in the limit of instantaneous steps ($k_2 \rightarrow \infty$ and $l = v/k_2$), and the original \citet{Einstein1937} model in the limit of instantaneous steps without burial.
These reductions demonstrate that the majority of recent bedload diffusion models, whether developed from Exner-type equations \citep{Wu2019,Pelosi2014,Pelosi2016} or advection-diffusion equations \citep{Lisle1998,Lajeunesse2017}, can be viewed equivalently as continuous-time random walks applied to individual bedload trajectories.
Within random walk theory, sophisticated descriptions of transport with variable velocities \citep{Zaburdaev2008,Masoliver1994}, correlated motions \citep{Escaff2018,Vicsek2012a}, and anomalous diffusion \citep{Masoliver2016,Fa2014,Metzler2014} have been developed.
Meanwhile, in bedload transport research, variable velocities \citep{Lajeunesse2010,Furbish2012,Heyman2016}, correlated motions \citep{Heyman2014,Lee2018,Saletti2020}, and anomalous diffusion \citep{Fathel2016,Bradley2017,Schumer2009} constitute open research issues.
We believe further developing the linkage between existing bedload models and random walk concepts could rapidly progress our understanding.

@@

>>>>

\documentclass[grl]{agujournal2018}
%\documentclass[]{agujournal2018}
\usepackage{apacite}
\usepackage{natbib}

\usepackage{wrapfig}
\usepackage{url}
\usepackage{lineno}
\usepackage{trackchanges}
\soulregister\citet7
\soulregister\citep7
\soulregister\ref7
\addeditor{JKP}

\draftfalse
\journalname{Geophysical Research Letters}

%custom packages
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{comment}
\usepackage{booktabs}
% custom commands
\newcommand\be{\begin{equation}}
\newcommand\ee{\end{equation}} 
\newcommand\bra{\langle}
\newcommand\ket{\rangle}
\newcommand\om{\omega}
\newcommand\tom{\tilde{\omega}}
\newcommand\tg{\tilde{g}}
\newcommand\tp{\tilde{p}}
\newcommand\tG{\tilde{G}}
\newcommand\El{\mathcal{L}}
%\usepackage{layouts}
%\printinunitsof{in}\prntlen{\textwidth} % check scales in the document
\linenumbers

% THIS IS THE LINENO PATCH TO WORK WITH AMSMATH
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
	\expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
	\expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
	\renewenvironment{#1}%
	{\linenomath\csname old#1\endcsname}%
	{\csname oldend#1\endcsname\endlinenomath}}% 
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
	\patchAmsMathEnvironmentForLineno{#1}%
	\patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
	\patchBothAmsMathEnvironmentsForLineno{equation}%
	\patchBothAmsMathEnvironmentsForLineno{align}%
	\patchBothAmsMathEnvironmentsForLineno{flalign}%
	\patchBothAmsMathEnvironmentsForLineno{alignat}%
	\patchBothAmsMathEnvironmentsForLineno{gather}%
	\patchBothAmsMathEnvironmentsForLineno{multline}%
}
% the patch comes from http://phaseportrait.blogspot.com/2007/08/lineno-and-amsmath-compatibility.html

\begin{document}

\title{Back to Einstein: burial-induced three-range diffusion in fluvial sediment transport}

\authors{J. Kevin Pierce \affil{1}and Marwan A. Hassan\affil{1}}
\affiliation{1}{Department of Geography \\University of British Columbia}
\correspondingauthor{Kevin Pierce}{kpierce@alumni.ubc.ca}

\begin{keypoints}
\item Random walk model describes coarse gravel tracers spreading out within a river as they gradually become buried
\item Interchange of grains between motion, surface rest, and burial states generates three diffusion ranges as the observation time increases
\item Sediment burial dominates long-time properties and exhumation may develop a fourth ``geomorphic'' range of tracer diffusion
\end{keypoints}

\begin{abstract}
Individual grains move through gravel bed rivers in cycles of motion and rest, so tracer grains diffuse as they transport downstream. Tracer experiments demonstrate at least three diffusion ranges as observation times increase, with different spreading rates in each range. Up till now, the generating processes of these ranges remains unclear. In this work, we develop a random walk model of individual bedload trajectories including motion, rest, and burial processes. The model describes three bedload diffusion ranges that terminate in a fourth non-diffusive range when all tracers become buried. Using the model, we attribute three-range tracer diffusion to the interplay between motion, rest, and burial processes, and we relate the multi-range diffusion characteristics to the timescales of these processes.
\end{abstract}

\section{Introduction}

@@

\section{Bedload trajectories as a multi-state random walk}
\label{sec:model}
\subsection{Model assumptions}
\label{sec:assumptions}
We construct a three-state random walk where the states are motion, surface rest, and burial, and we label these states as $i=2$ (motion), $i=1$ (rest), and $i=0$ (burial).
Our target is the probability distribution $p(x,t)$ to find a grain at position $x$ and time $t$ if we know it started with the initial distribution $p(x,0)=\delta(x)$.
We characterize times spent moving or resting on the surface by exponential distributions $\psi_2(t)&=k_2e^{-k_2 t}$ and $\psi_1(t) &= k_1e^{-k_1t}$, since numerous experiments show thin-tailed distributions for these quantities \citep{Fathel2015,Roseberry2012,Einstein1937,Ancey2006,Martin2012}. We expect our conclusions will not be contingent on the specific distributions chosen, since all thin-tailed distributions provide similar diffusion characteristics in random walks \citep{Weiss1994,Weeks1998}.
We consider grains in motion to have characteristic velocity $v$ \citep{Lisle1998,Lajeunesse2018}, and we model burial as long lasting enough to be effectively permanent \citep{Wu2019}, with grains resting on the surface having a probability per unit time $\kappa$ to become buried, meaning $\Phi(t) = e^{-\kappa t}$ represents the probability that a grain is not buried after resting for a time $t$, while $1-\Phi(t)$ represents the probability that it is buried.
We specify the initial conditions with probabilities $\theta_1$ and $\theta_2$ to be in rest and motion at $t=0$, and we require $\theta_1+\theta_2=1$ for normalization.

@@

\begin{figure}[t]
	\includegraphics[width=\linewidth,keepaspectratio]{pdf-plot.png}
	\caption{Joint distributions for a grain to be at position $x$ at time $t$ are displayed for the choice $k_1=0.1$, $k_2=1.0$, $v=2.0$. Grains are considered initially at rest ($\theta_1=1$, $\theta_2=0$). The solid lines are the analytical distribution in equation (\ref{eq:pdf}), while the points are numerically simulated, showing the correctness of our derivations. Colors pertain to different times. Units are unspecified, since we aim to demonstrate the general characteristics of $p(x,t)$. Panel (a) shows the case $\kappa=0$ -- no burial. In this case, the joint distribution tends toward Gaussian at large times \citep{Einstein1937,Lisle1998}. Panel (b) shows the case when grains have rate $\kappa = 0.01$ to become buried while resting. Because of burial, the joint distribution tends toward a more uniform distribution than Gaussian.
	\label{fig:pdfs}}
\end{figure}

@@

\begin{figure}[h!]	
	\includegraphics[width=\linewidth,keepaspectratio]{diffusion.png}
	\caption{Panel (a) sketches conceptual trajectories of three grains, while panel (b) depicts the variance (\ref{eq:var}) with mean motion time $1.5$ s, resting time $30.0$ s, and movement velocity $0.1$ m/s -- values comparable to laboratory experiments transporting small ($5$ mm) gravels \citep{Lajeunesse2010,Martin2012}. The burial timescale is $7200.0$s (two hours), and grains start from rest ($\theta_1=1$). The solid line is equation (\ref{eq:var}), and the points are numerically simulated. Panel (b) demonstrates four distinct scaling ranges of $\sigma_x^2$: local, intermediate, global, and geomorphic. The first three are diffusive. Three crossover times $\tau_L$, $\tau_I$, and $\tau_G$ divide the ranges. Within each range, a slope key demonstrates the scaling $\sigma_x^2 \propto t^\gamma$. Panel (a) demonstrates that different mixtures of motion, rest, and burial states generate the ranges. At local timescales, grains usually either rest or move; at intermediate timescales, they transition between rest and motion; at global timescales, they transition between rest, motion, and burial; and at geomorphic timescales, all grains bury. Additional slope keys in the local and global ranges of panel (b) illustrate the effect of initial conditions and rest/burial timescales on the diffusion, while the additional slope key within the geomorphic range demonstrates the expected scaling when burial is not permanent, as we discuss in section \ref{sec:discussion}.
}
	\label{fig:var}
\end{figure}

@@

The local and intermediate range diffusion characteristics resulting from our model correspond closely to the original Nikora et al. concepts, while our global range has a different origin than Nikora et al. envisioned.
\citet{Nikora2001a} explained that local diffusion results from the non-fractal (smooth) characteristics of bedload trajectories between subsequent interactions with the bed,  while intermediate diffusion results from the fractal (rough) characteristics of bedload trajectories after many interactions with the bed.
Our model represents these conclusions: non-fractal (and super-diffusive) bedload trajectories exist on timescales short enough that each grain is either resting or moving, while fractal (and normal-diffusive) bedload trajectories exist on timescales when grains are actively switching between motion and rest states.
We conclude that local and intermediate ranges stem from the interplay between motion and rest timescales, as demonstrated by earlier two-state random walk models \citep{Lisle1998,Lajeunesse2018} and by all Newtonian models that develop sequences of motions and rests \citep{Nikora2001a, Bialik2012}, even those including heavy-tailed rests \citep{Fan2016}.

@@

The analytical solution of bedload diffusion in equation (\ref{eq:var}) reduces exactly to the analytical solutions of the \citet{Lisle1998} and \citet{Lajeunesse2018} models in the limit without burial ($\kappa \rightarrow 0$), the \citet{Wu2019} model in the limit of instantaneous steps ($k_2 \rightarrow \infty$ and $l = v/k_2$), and the original \citet{Einstein1937} model in the limit of instantaneous steps without burial.
These reductions demonstrate that the majority of recent bedload diffusion models, whether developed from Exner-type equations \citep{Wu2019,Pelosi2014,Pelosi2016} or advection-diffusion equations \citep{Lisle1998,Lajeunesse2018}, can be viewed equivalently as continuous-time random walks applied to individual bedload trajectories.
Within random walk theory, sophisticated descriptions of transport with variable velocities \citep{Zaburdaev2008,Masoliver1994}, correlated motions \citep{Escaff2018,Vicsek2012a}, and anomalous diffusion \citep{Masoliver2016,Fa2014,Metzler2014} have been developed.
Meanwhile, in bedload transport research, variable velocities \citep{Lajeunesse2010,Furbish2012,Heyman2016}, correlated motions \citep{Heyman2014,Lee2018,Saletti2020}, and anomalous diffusion \citep{Fathel2016,Bradley2017,Schumer2009} constitute open research issues.
We believe further developing the linkage between existing bedload models and random walk concepts could rapidly progress our understanding.

@@

\appendix

\acknowledgments
K. Pierce acknowledges thoughtful exchanges with Eduardo Daly and Eric Weeks during the early stages of this work. Both authors would like to thank Matteo Saletti, Conor McDowell, Shawn Chartrand, and Will Booker for helpful discussions. M. Hassan is supported by an NSERC Discovery grant. The Python simulation code is freely available at \sloppy
\url{https://zenodo.org/record/3659291#.Xj3p7XVKjIU}.

\bibliography{biblio.bib}
\end{document}

### 2021-06-24 pref.tex
got one chapter in. some issues.

\chapter{Acknowledgments}

Thanks Leo Golubovic, Mindy Saunders, Calisa and Jim Pierce, Kim and Kelsey Pierce, Johnathan Ramey, Linda Mendez, Tex Wood, Charles Wood, and all of the Appalachian intelligensia who paved my way. Tyler cannon. . .

Shawn Chartrand, Conor McDowell, Matteo Saletti, Will Booker, David Adams, Leo King, Tobias Mueller, Nisreen Ghazi,

Adam Payne, Adam Collins, Megat Denney, Andrew Rice, Evan Graber, Scott Ferris, Craig Tenney, and the rest of the WVU physics lounge family.

Andy Osadetz, Josh Plankeel,

My partner Mary for putting up with the late nights.

>>>>

\chapter{Preface}

A preface should be written indicating what is to come and what the point of this whole thing is.
